### 2020-07-29 WorkingScriptJuly29.Rmd
Counted my sample sizes

---
title: "Working Script"
output: html_notebook
---

This script is just to play around with the data while I'm still collecting it.

## Load in packages

```{r}
library(dplyr)
library(tidyr)
```

## Load in the data

```{r}
data <- read.csv("~/Desktop/MsCcHSFeedJuly29.csv")
```

## Counting sample size

```{r}
count(data[data$food == "TB" & data$para == "yes" & data$HS == "yes", ])
count(data[data$food == "TB" & data$para == "yes" & data$HS == "no", ])
count(data[data$food == "TB" & data$para == "no" & data$HS == "yes", ])
count(data[data$food == "TB" & data$para == "no" & data$HS == "no", ])
```

```{r}
count(data[data$food == "DC" & data$para == "yes" & data$HS == "yes", ])
count(data[data$food == "DC" & data$para == "yes" & data$HS == "no", ])
count(data[data$food == "DC" & data$para == "no" & data$HS == "yes", ])
count(data[data$food == "DC" & data$para == "no" & data$HS == "no", ])
```

## Check how many have survived

```{r}
data <- data[data$survived == "yes", ]
```

```{r}
count(data[data$food == "TB" & data$para == "yes" & data$HS == "yes", ])
count(data[data$food == "TB" & data$para == "yes" & data$HS == "no", ])
count(data[data$food == "TB" & data$para == "no" & data$HS == "yes", ])
count(data[data$food == "TB" & data$para == "no" & data$HS == "no", ])
```

```{r}
count(data[data$food == "DC" & data$para == "yes" & data$HS == "yes", ])
count(data[data$food == "DC" & data$para == "yes" & data$HS == "no", ])
count(data[data$food == "DC" & data$para == "no" & data$HS == "yes", ])
count(data[data$food == "DC" & data$para == "no" & data$HS == "no", ])
```