### 2020-09-01 WorkingScript.Rmd
Figured out how to get the total leaf area eaten for each individual.
Preliminary model.

---
title: "Working Script"
output: html_notebook
---

This script is just to play around with the data while I'm still collecting it.

## Load in packages

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
```

## Load in the data

```{r}
data <- read.csv("~/Desktop/MsCcHSFeedAug30.csv")
```

## Counting sample size

```{r}
count(data[data$food == "TB" & data$para == "yes" & data$HS == "yes", ])
count(data[data$food == "TB" & data$para == "yes" & data$HS == "no", ])
count(data[data$food == "TB" & data$para == "no" & data$HS == "yes", ])
count(data[data$food == "TB" & data$para == "no" & data$HS == "no", ])
```

```{r}
count(data[data$food == "DC" & data$para == "yes" & data$HS == "yes", ])
count(data[data$food == "DC" & data$para == "yes" & data$HS == "no", ])
count(data[data$food == "DC" & data$para == "no" & data$HS == "yes", ])
count(data[data$food == "DC" & data$para == "no" & data$HS == "no", ])
```

## Check how many have survived

```{r}
data <- data[data$survived == "yes", ]
```

```{r}
count(data[data$food == "TB" & data$para == "yes" & data$HS == "yes", ])
count(data[data$food == "TB" & data$para == "yes" & data$HS == "no", ])
count(data[data$food == "TB" & data$para == "no" & data$HS == "yes", ])
count(data[data$food == "TB" & data$para == "no" & data$HS == "no", ])
```

```{r}
count(data[data$food == "DC" & data$para == "yes" & data$HS == "yes", ])
count(data[data$food == "DC" & data$para == "yes" & data$HS == "no", ])
count(data[data$food == "DC" & data$para == "no" & data$HS == "yes", ])
count(data[data$food == "DC" & data$para == "no" & data$HS == "no", ])
```

## Load in the leaf area data

```{r}
scan <- read.csv("~/Desktop/MsCcHSFeedLeafScanAug29.csv", header = T)
```

## Write a for-loop for getting the amounts eaten

0. Create output file 
1. Combine Leaf ID and cat ID into one ID 
2. Parse by unique  ID
3. Subtract min measurement from max measurement
4. Record with comb ID
5. Separate out catID and leaf ID
6. Create new output structure
7. Sum for each cat ID 
8. Combine with main dataset as a column

```{r}
#1
scan$comb <- paste(scan$Caterpillar.ID, "_", scan$Leaf.ID)

#0
unique <- unique(scan$comb)
consumed <- as.data.frame(unique)
consumed$area <- 0

#2-4
for (i in 1:length(consumed$unique)){
    subset <- filter(scan, scan$comb == consumed[i, 1])
    value <- max(subset$Leaf.Area) - min(subset$Leaf.Area)
    consumed[i, 2] <- value
}

#5
consumed <- consumed %>%
  separate(unique, c("catID", "leafID"), "_")

#6
ID <- unique(consumed$catID)
consumedsum <- as.data.frame(ID)
consumedsum$totalarea <- 0

#7
for (i in 1:length(consumedsum$ID)){
  subset2 <- filter(consumed, consumed$catID == consumedsum[i, 1])
  value2 <- sum(subset2$area)
  consumedsum[i, 2] <- value2
}

```

Step 8 proved to be more complicated, so it's listed below...

```{r}
consumedsum$ID <- as.numeric(consumedsum$ID)

consumedsum <- consumedsum[consumedsum$ID != 707, ]
consumedsum <- consumedsum[consumedsum$ID != 1816, ]

consumedsum <- consumedsum[consumedsum$ID %in% data$ID, ]

data <- data[data$ID %in% consumedsum$ID, ]

consumedsum <- consumedsum[order(consumedsum$ID), ]

data$area <- consumedsum$totalarea
```

YAY!!!

Now, let's run some preliminary analyses:

```{r}
mod1 <- lmer(area ~ food + (1|bin), data = data)

summary(mod1)
```

```{r}
dataDC <- data[data$food == "DC", ]
dataTB <- data[data$food == "TB", ]

t.test(dataDC$area, dataTB$area)
```