### 2020-12-07 ASNCode.Rmd
Changed some colors, added some graphs

---
title: "Code for Graphs for ASN presentation"
output: html_notebook
---
This document contains the code for the graphs I'll include in my 2021 ASN presentation, as well as model results I'll present.

As such, it doesn't contain all models or all graphs - see WorkingScript.Rmd for a stream-of-consciousness view of my work.

## Load in packages

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lubridate)
library(viridisLite)
```

## Load in data

```{r}
data <- read.csv("~/Desktop/MsCcHSFeedNov30modfinal.csv")
scan <- read.csv("~/Desktop/MsCcHSFeedLeafScanFinal.csv", header = T)
```

## Manipulate data into correct forms

```{r}
data <- data[data$fate != "dead" & data$fate != "quink" & data$fate != "wander*" & data$fate != "?", ]
```

```{r}
#1
scan$comb <- paste(scan$Caterpillar.ID, "_", scan$Leaf.ID)

#0
unique <- unique(scan$comb)
consumed <- as.data.frame(unique)
consumed$area <- 0

#2-4
for (i in 1:length(consumed$unique)){
    subset <- filter(scan, scan$comb == consumed[i, 1])
    value <- max(subset$Leaf.Area) - min(subset$Leaf.Area)
    consumed[i, 2] <- value
}

#5
consumed <- consumed %>%
  separate(unique, c("catID", "leafID"), "_")

#6
ID <- unique(consumed$catID)
consumedsum <- as.data.frame(ID)
consumedsum$totalarea <- 0

#7
for (i in 1:length(consumedsum$ID)){
  subset2 <- filter(consumed, consumed$catID == consumedsum[i, 1])
  value2 <- sum(subset2$area)
  consumedsum[i, 2] <- value2
}

#8
consumedsum$ID <- as.numeric(consumedsum$ID)

consumedsum <- consumedsum[consumedsum$ID %in% data$ID, ]

data <- data[data$ID %in% consumedsum$ID, ]

consumedsum <- consumedsum[order(consumedsum$ID), ]

data$area <- consumedsum$totalarea

```

```{r}
# Internal process of the loop to convert days to dates
expt_day <- function(start.date, date){
  date <- mdy(date)
  start <- mdy(start.date)
  diff <- yday(date) - yday(start)

  return(diff)
}

# Code for the loop itself
loop <- function(start.date, date.list){
  days <- rep(NA, length(date.list))
  for (i in 1:(length(days))){
    days[i] <- expt_day(start.date = start.date, date = date.list[i])
  }
  return(days)
}
```

```{r}
data$day.hatch <- loop(start.date = "07/12/2020", date.list = data$date.hatch)
data$day.max <- loop(start.date = "07/12/2020", date.list = data$date.max)

data$ttmax <- data$day.max - data$day.hatch
```

## Graphs

```{r}
TBdata <- data[data$food == "TB", ]
DCdata <- data[data$food == "DC", ]
```

Four colors I'm using:

emerged: #66CB5D
wander: #21908D
wowe: #3E4F8A
wowe mongo: #450D54

### Wasp fate

### Maximum mass gained and age at that point

```{r}
ggplot(aes(x = ttmax, y = mass.max, color = fate), data = DCdata) +
  theme_bw() + 
  geom_point(alpha = 0.8) + 
  scale_color_manual(values = c("#66CB5D", "#21908D", "#3E4F8A", "#450D54")) + 
  labs(x = "Development time at maximum mass (days)", y = "Maximum mass achieved (mg)", 
       color = "Developmental Outcome")
```

```{r}
dataPHS <- data[data$treatment == "P HS", ]

ggplot(aes(x = ttmax, y = mass.max, color = fate), data = dataPHS) +
  theme_bw() + 
  geom_point(alpha = 0.8) + 
  scale_color_manual(values = c("#66CB5D", "#21908D", "#3E4F8A", "#450D54")) + 
  labs(x = "Development time at maximum mass (days)", y = "Maximum mass achieved (mg)", 
       color = "Developmental Outcome")
```

### Maximum mass compared to plant matter consumed

```{r}
ggplot(aes(x = area, y = mass.max), data = DCdata) + 
  theme_bw() + 
  geom_point(aes(color = fate)) + 
  scale_color_manual(values = c("#66CB5D", "#21908D", "#3E4F8A", "#450D54")) +
  labs(x = "Leaf surface area consumed (cm2)", y = "Maximum mass achieved (mg)",
       color = "Developmental Outcome")
```
```{r}

```