### 2022-01-01 ArchivedCode/MsCcHSFeedOutlineNov10.Rmd

---
title: "Samples Sizes - MsCcHSFeed"
output: html_notebook
---

```{r}
library(ggplot2)
library(tidyverse)
library(Rmisc)
```

```{r}
data <- read.csv("~/Desktop/MsCcHSFeedJan1MOD.csv")
scan <- read.csv("~/Desktop/MsCcHSFeedLeafScanFinal.csv", header = T)
```

## Manipulate data into correct forms

```{r}
data <- data[data$fate != "dead" & data$fate != "quink" & data$fate != "wander*" & data$fate != "?", ]
data <- data[complete.cases(data[, 1]), ]
```

## Count stuff

```{r}
count(data[data$food == "TB" & data$para == "yes" & data$HS == "yes", ])
count(data[data$food == "TB" & data$para == "yes" & data$HS == "no", ])
count(data[data$food == "TB" & data$para == "no" & data$HS == "yes", ])
count(data[data$food == "TB" & data$para == "no" & data$HS == "no", ])
```

```{r}
count(data[data$food == "DC" & data$para == "yes" & data$HS == "yes", ])
count(data[data$food == "DC" & data$para == "yes" & data$HS == "no", ])
count(data[data$food == "DC" & data$para == "no" & data$HS == "yes", ])
count(data[data$food == "DC" & data$para == "no" & data$HS == "no", ])
```

## Make graph of outcomes

```{r}
data$treatment <- factor(data$treatment, levels = c("NP NHS", "NP HS", "P NHS", "P HS"))
```

```{r}
bars <- ggplot(data = data, aes(x = treatment, fill = fate)) + 
  theme_bw() + 
  geom_bar(position = "fill") + 
  labs(x = "Treatment group", y = "Proportion of indiviudals", fill = "Developmental Outcome")

bars
```

```{r}
bars2 <- ggplot(data = data, aes(x = treatment, fill = fate)) + 
  theme_bw() + 
  geom_bar(position = "stack") + 
  facet_wrap(~food) + 
  labs(x = "Treatment group", y = "Number of indiviudals", fill = "Developmental Outcome")

bars2

ggsave("~/Desktop/outcomes.png", plot = bars2, width = 8, height = 5)
```

## This time, keep the dead

```{r}
datadead <- read.csv("~/Desktop/MsCcHSFeedJan1MOD.csv")
datadead <- datadead[datadead$fate != "quink" & datadead$fate != "wander*" & datadead$fate != "?", ]
datadead <- datadead[complete.cases(datadead[, 1]), ]
datadead$treatment <- factor(datadead$treatment, levels = c("NP NHS", "NP HS", "P NHS", "P HS"))
```

```{r}
barsdead <- ggplot(data = datadead, aes(x = treatment, fill = fate)) + 
  theme_bw() + 
  geom_bar(position = "fill") + 
  facet_wrap(~food) + 
  labs(x = "Treatment group", y = "Number of indiviudals", fill = "Developmental Outcome") + 
  scale_fill_manual(values = c("#D81B60", "orange", "lime green", "#1E88E5", "purple"))

barsdead

ggsave("~/Desktop/outcomesdead.png", plot = barsdead, width = 8, height = 5)
```

# Growth trajectories of NP NHS vs NP HS

## Manipulate the data

```{r}
# Internal process of the loop to convert days to dates
expt_day <- function(start.date, date){
  date <- mdy(date)
  start <- mdy(start.date)
  diff <- yday(date) - yday(start)

  return(diff)
}

# Code for the loop itself
loop <- function(start.date, date.list){
  days <- rep(NA, length(date.list))
  for (i in 1:(length(days))){
    days[i] <- expt_day(start.date = start.date, date = date.list[i])
  }
  return(days)
}
```

```{r}
data$day.hatch <- loop(start.date = "07/12/2020", date.list = data$date.hatch)
data$day.3rd <- loop(start.date = "07/12/2020", date.list = data$date.3rd)
data$day.4th <- loop(start.date = "07/12/2020", date.list = data$date.4th)
data$day.5th <- loop(start.date = "07/12/2020", date.list = data$date.5th)
data$day.w <- loop(start.date = "07/12/2020", date.list = data$date.w)

data$tthatch <- 0
data$tt3rd <- data$day.3rd - data$day.hatch
data$tt4th <- data$day.4th - data$day.hatch
data$tt5th <- data$day.5th - data$day.hatch
data$ttw <- data$day.w - data$day.hatch
```

```{r}
data$ln.hatch <- log(data$mass.hatch)
data$ln.3rd <- log(data$mass.3rd)
data$ln.4th <- log(data$mass.4th)
data$ln.5th <- log(data$mass.5th)
data$ln.w <- log(data$mass.w)
```

```{r}
dataslim <- data[data$treatment == "NP NHS" | data$treatment == "NP HS", ] %>%
  select(ID, food, fate, treatment, tthatch, ln.hatch, tt3rd, ln.3rd, tt4th, ln.4th, tt5th, ln.5th, ttw, ln.w)

```

```{r}
masslong <- dataslim %>%
  gather(instar, mass, ln.hatch, ln.3rd, ln.4th, ln.5th, ln.w)
masslong$instar <- gsub("ln.", "", masslong$instar)

agelong <- dataslim %>%
  gather(instar, age, tthatch, tt3rd, tt4th, tt5th, ttw)
agelong$instar <- gsub("tt", "", agelong$instar)

masslong <- masslong %>% 
  select(ID, treatment, food, instar, mass)
agelong <- agelong %>% 
  select(ID, treatment, food, instar, age)

datalong<-merge(masslong, agelong, by=c("ID", "treatment", "food", "instar"))

datalong$instar <- factor(datalong$instar, levels = c("hatch", "3rd", "4th", "5th", "w"))

datalong$newtreatment <- paste(datalong$treatment, datalong$food)
datalong$newtreatment <- factor(datalong$newtreatment, levels = c("NP NHS DC", "NP NHS TB", "NP HS DC", "NP HS TB"))
```

```{r}

#calculate means and variation

#constructing summarySE for age and mass to make mean plot

masssum <- summarySE(datalong, measurevar = "mass",
                  groupvars = c("newtreatment", "instar"),
                  na.rm=TRUE)
masssum

agesum <- summarySE(datalong, measurevar = "age",
                   groupvars = c("newtreatment", "instar"),
                   na.rm = TRUE)
agesum

#combine into one data frame

masssum$age <- agesum[, 4]
masssum$age.se <- agesum[, 6]
masssum$age.ci <- agesum[, 7]
```

## Plot the data (with CIs)

```{r}
meanplot <- ggplot(masssum, aes(x = age, y = mass, group = newtreatment, color = newtreatment, shape= newtreatment)) + 
  theme_bw() + 
  geom_point(size = 4) + 
  geom_line(aes(linetype = newtreatment), size = 1) + 
  geom_errorbar(aes(ymin = mass-(ci/2), ymax = mass + (ci/2)),
                width = .3, size = 1) + 
  geom_errorbarh(aes(xmin = age - (age.ci/2), xmax = age + (age.ci/2)),
                 height = .3, size = 1) + 
  scale_color_manual(values = c("#D81B60","#1E88E5", "#D81B60","#1E88E5" )) +
  scale_shape_manual(values = c(1, 1, 2, 2)) + 
  scale_linetype_manual(values = c("dashed", "dashed", "solid", "solid")) + 
  labs(x = "Total developmental age (days)", y = "Total mass gained (ln)")

ggsave("meanplot.png", plot = meanplot, width = 8, height = 5)

meanplot
```