### 2022-01-18 MsCcHSFeedOutlineDec15.Rmd
Added new wasp survival figure and analyses, relabeled other figure numbers!

---
title: "MsCcHSFeed Outline"
output: html_notebook
---

```{r}
library(ggplot2)
library(tidyverse)
library(Rmisc)
library(lubridate)
library(lmerTest)
```

```{r}
data <- read.csv("~/Desktop/MsCcHSFeedJan1MOD.csv")
scan <- read.csv("~/Desktop/MsCcHSFeedLeafScanFinal.csv", header = T)
```

```{r}
data <- data[data$fate != "dead" & data$fate != "quink" & data$fate != "wander*" & data$fate != "?", ]
data <- data[complete.cases(data[, 1]), ]
data$fate <- factor(data$fate, levels = c("wander", "em", "wowe", "wowe mongo"))

```

```{r}
datadead <- read.csv("~/Desktop/MsCcHSFeedJan1MOD.csv")
datadead <- datadead[datadead$fate != "quink" & datadead$fate != "wander*" & datadead$fate != "?", ]
datadead <- datadead[complete.cases(datadead[, 1]), ]
datadead$treatment <- factor(datadead$treatment, levels = c("NP NHS", "NP HS", "P NHS", "P HS"))
```

# Fig 1: Counts

```{r}
count(datadead[datadead$food == "TB" & datadead$para == "yes" & datadead$HS == "yes", ])
count(datadead[datadead$food == "TB" & datadead$para == "yes" & datadead$HS == "no", ])
count(datadead[datadead$food == "TB" & datadead$para == "no" & datadead$HS == "yes", ])
count(datadead[datadead$food == "TB" & datadead$para == "no" & datadead$HS == "no", ])
```

```{r}
count(datadead[datadead$food == "DC" & datadead$para == "yes" & datadead$HS == "yes", ])
count(datadead[datadead$food == "DC" & datadead$para == "yes" & datadead$HS == "no", ])
count(datadead[datadead$food == "DC" & datadead$para == "no" & datadead$HS == "yes", ])
count(datadead[datadead$food == "DC" & datadead$para == "no" & datadead$HS == "no", ])
```

# Fig 2: Wasp survival across plant types

```{r}
datawasp <- data[data$para == "yes", ] %>%
  select(ID, food, treatment, num.w.coc, num.w.larvae, num.w.eclos)

datawasp <- datawasp[complete.cases(datawasp), ]

datawasp$num.w.tot <- datawasp$num.w.coc + datawasp$num.w.larvae
datawasp$percent <- (datawasp$num.w.eclos) / (datawasp$num.w.tot)
```

```{r}
waspplot <- ggplot(data = datawasp, aes(x = food, y = percent)) +
  theme_bw() + 
  geom_boxplot() + 
  geom_jitter(aes(color = treatment), width = 0.1, height = 0) + 
  labs(x = "Hostplant", y = "Percent survival from emergence to eclosion", color = "Treatment")

waspplot

ggsave("~/Desktop/waspplot.png", plot = waspplot, width = 8, height = 5)

```

## Statistics

```{r}
waspbinom <- glm(percent ~ food, family = "binomial", weights = num.w.tot, data = datawasp)

summary(waspbinom)

anova(waspbinom, test = "Chisq")

```

```{r}
waspbinom2 <- glm(percent ~ food, family = "binomial", weights = num.w.tot, 
                  data = datawasp[datawasp$treatment == "P NHS", ])

summary(waspbinom2)

anova(waspbinom2, test = "Chisq")

```

```{r}
waspemerge <- lm(num.w.tot ~ food, data = datawasp)

summary(waspemerge)
```

```{r}
waspemergetreat <- lm(num.w.tot ~ treatment, data = datawasp)

summary(waspemergetreat)
```

# Fig 3: Violin plots of max mass

```{r}
data$treatment <- factor(data$treatment, levels = c("NP NHS", "NP HS", "P NHS", "P HS"))

dataviolin <- data
dataviolin$fate <- gsub(" mongo", "", dataviolin$fate)
dataviolin$fate <- factor(dataviolin$fate, levels = c("wander", "em", "wowe"))

violins <- ggplot(data = dataviolin, aes(x = treatment, y = mass.max)) + 
  geom_violin(fill = "light grey", color = "light grey") + 
  geom_jitter(aes(color = fate), width = 0.1, height = 0) + 
  theme_bw() + 
  labs(x = "Treatment Group", y  = "Maximum mass achieved (mg)",  color = "Developmental Outcome") + 
  scale_color_manual(values = c("#1E88E5", "lime green", "#D81B60"), 
                     labels = c("Wander", "WE", "WOWE"))

violins

ggsave("~/Desktop/violins2.png", violins, width = 8, height = 5)

```

# Fig 4: Developmental outcome proportions

## Make graph of outcomes

```{r}
data$treatment <- factor(data$treatment, levels = c("NP NHS", "NP HS", "P NHS", "P HS"))
datadead$fate <- factor(datadead$fate, levels = c("dead", "wander", "em", "wowe", "wowe mongo"))
```

## This time, keep the dead

```{r}
barsdead <- ggplot(data = datadead, aes(x = treatment, fill = fate)) + 
  theme_bw() + 
  geom_bar(position = "fill") + 
  facet_wrap(~food) + 
  labs(x = "Treatment group", y = "Proportion of indiviudals", fill = "Developmental Outcome") + 
  scale_fill_manual(values = c("purple", "#1E88E5", "lime green", "orange", "#D81B60"),
                    labels = c("Died", "Wander", "WE", "WOWE", "WOWE mongo"))

barsdead

ggsave("~/Desktop/outcomesdead.png", plot = barsdead, width = 8, height = 5)
```

# Fig 5: Maximum mass and age

## Manipulate the data
```{r}
# Internal process of the loop to convert days to dates
expt_day <- function(start.date, date){
  date <- mdy(date)
  start <- mdy(start.date)
  diff <- yday(date) - yday(start)

  return(diff)
}

# Code for the loop itself
loop <- function(start.date, date.list){
  days <- rep(NA, length(date.list))
  for (i in 1:(length(days))){
    days[i] <- expt_day(start.date = start.date, date = date.list[i])
  }
  return(days)
}
```

```{r}
data$day.hatch <- loop(start.date = "07/12/2020", date.list = data$date.hatch)
data$day.max <- loop(start.date = "07/12/2020", date.list = data$date.max)
data$ttmax <- data$day.max - data$day.hatch

data$mass.max <- data$mass.max / 1000
```

```{r}
#1
scan$comb <- paste(scan$Caterpillar.ID, "_", scan$Leaf.ID)

#0
unique <- unique(scan$comb)
consumed <- as.data.frame(unique)
consumed$area <- 0

#2-4
for (i in 1:length(consumed$unique)){
    subset <- filter(scan, scan$comb == consumed[i, 1])
    value <- max(subset$Leaf.Area) - min(subset$Leaf.Area)
    consumed[i, 2] <- value
}

#5
consumed <- consumed %>%
  separate(unique, c("catID", "leafID"), "_")

#6
ID <- unique(consumed$catID)
consumedsum <- as.data.frame(ID)
consumedsum$totalarea <- 0

#7
for (i in 1:length(consumedsum$ID)){
  subset2 <- filter(consumed, consumed$catID == consumedsum[i, 1])
  value2 <- sum(subset2$area)
  consumedsum[i, 2] <- value2
}

#8
consumedsum$ID <- as.numeric(consumedsum$ID)

consumedsum <- consumedsum[consumedsum$ID %in% data$ID, ]

data <- data[data$ID %in% consumedsum$ID, ]

consumedsum <- consumedsum[order(consumedsum$ID), ]

data$area <- consumedsum$totalarea

data$area <- data$area / 10000

```

```{r}
datalong5 <- data %>%
  select(ID, food, fate, mass.max, ttmax, area)

meantime2 <- summarySE(datalong5, measurevar = "ttmax", groupvars = c("food", "fate"), na.rm = TRUE)

meanmass2 <- summarySE(datalong5, measurevar = "mass.max", groupvars = c("food", "fate"), na.rm = TRUE)

meanarea2 <- summarySE(datalong5, measurevar = "area", groupvars = c("food", "fate"), na.rm = TRUE)
```

### Group them into one dataframe

```{r}
datameans2 <- meantime2 %>%
  select(food, fate, ttmax, se)

datameans2$mass.max <- meanmass2$mass.max
datameans2$semass <- meanmass2$se
datameans2$area <- meanarea2$area
datameans2$arease <- meanarea2$se

```

```{r}
newmeans <- ggplot(data = datalong5, aes(x = ttmax, y = mass.max, color = fate, shape = food)) + 
  theme_bw() + 
  xlim(15, 45) +
  geom_point(alpha = 0.4) + 
  scale_shape_manual(values = c(1, 19)) + 
  scale_color_manual(values = c("#1E88E5", "lime green", "orange", "#D81B60"), 
                     labels = c("Wander", "WE", "WOWE", "WOWE mongo")) + 
  geom_point(aes(x = ttmax, y = mass.max, color = fate, shape = food), data = datameans2, size = 4) + 
  geom_errorbar(data = datameans2, aes(ymin = mass.max - semass, ymax = mass.max + semass), size = .5) + 
  geom_errorbarh(data = datameans2, aes(xmin = ttmax - se, xmax = ttmax + se), size = .5) + 
  guides(size = FALSE) + 
  labs(x = "Development time at maximum mass (days)", y = "Maximum mass achieved (g)", 
       shape = "Hostplant Consumed", color = "Developmental Outcome")

newmeans

ggsave("~/Desktop/newmeans.png", newmeans, width = 8, height = 5)
```

## Stats

### Maximum age differences across treatment and outcome

```{r}
maxage <- lmer(ttmax ~ fate*food + (1|bin), data = data)

anova(maxage)
```

```{r}
maxage2 <- lmer(ttmax ~ para*HS*food + (1|bin), data = data)

anova(maxage2)
```

```{r}
AIC(maxage, maxage2)
```

### Maximum mass

```{r}
maxmass <- lmer(mass.max ~ fate*food + (1|bin), data = data)

anova(maxmass)
```

```{r}
maxmass2 <- lmer(mass.max ~ para*HS*food + (1|bin), data = data)

anova(maxmass2)
```

```{r}
AIC(maxmass, maxmass2)
```

# Fig 6: Leaf matter consumed vs. max mass

```{r}
newarea <- ggplot(data = datalong5, aes(x = area, y = mass.max, color = fate, shape = food)) + 
  theme_bw() + 
  geom_point(alpha = 0.4) + 
  scale_shape_manual(values = c(1, 19)) + 
  scale_color_manual(values = c("#1E88E5", "lime green", "orange", "#D81B60"), 
                     labels = c("Wander", "WE", "WOWE", "WOWE mongo")) + 
  ylim(0, 20) + 
  xlim(0, .6) +
  geom_point(aes(x = area, y = mass.max, color = fate, shape = food), data = datameans2, size = 4) + 
  geom_errorbar(data = datameans2, aes(ymin = mass.max - semass, ymax = mass.max + semass), size = .5) + 
  geom_errorbarh(data = datameans2, aes(xmin = area - arease, xmax = area + arease), size = .5) + 
  guides(size = FALSE) + 
  labs(x = "Hostplant leaf area consumed (m^2)", y = "Maximum mass achieved (g)", 
       shape = "Hostplant Consumed", color = "Developmental Outcome")

newarea

ggsave("~/Desktop/newarea.png", newarea, width = 8, height = 5)
```

## Stats

```{r}
areatrt <- lmer(area ~ fate*food + (1|bin), data = data)

anova(areatrt)
```

```{r}
areatrt2 <- lmer(area ~ para*HS*food + (1|bin), data = data)

anova(areatrt2)
```

```{r}
AIC(areatrt, areatrt2)
```

```{r}
anova(areatrt, areatrt2)
```

# Additional analyses

## Binomial of deaths between NP and P groups

```{r}
binomdeath <- datadead %>%
  select(ID, food, para, survived)

binomdeath$survived <- gsub("yes", 1, binomdeath$survived)
binomdeath$survived <- gsub("no", 0, binomdeath$survived)

binomdeath$survived <- as.numeric(binomdeath$survived)
```

### Run stats

```{r}
death <- glm(survived ~ food*para, data = binomdeath, family = "binomial")

summary(death)
```

```{r}
death2 <- glm(survived ~ para, data = binomdeath, family = "binomial")

summary(death2)
```

## Binomial of WOWE mongo outcomes

```{r}
binomWOWE <- data[data$treatment == "P HS", ] %>%
  select(ID, food, fate)

binomWOWE$fate <- gsub("wowe mongo", 1, binomWOWE$fate)
binomWOWE$fate <- gsub("wowe", 0, binomWOWE$fate)
binomWOWE$fate <- gsub("em", 0, binomWOWE$fate)
binomWOWE$fate <- gsub("wander", 0, binomWOWE$fate)

binomWOWE$fate <- as.numeric(binomWOWE$fate)
```

### Run stats

```{r}
wowe <- glm(fate ~ food, data = binomWOWE, family = "binomial")

summary(wowe)
```

## Binomial of WE outcomes

```{r}
binomWE <- data[data$treatment == "P HS", ] %>%
  select(ID, food, fate)

binomWE$fate <- gsub("wowe mongo", 0, binomWE$fate)
binomWE$fate <- gsub("wowe", 0, binomWE$fate)
binomWE$fate <- gsub("em", 1, binomWE$fate)
binomWE$fate <- gsub("wander", 0, binomWE$fate)

binomWE$fate <- as.numeric(binomWE$fate)
```

```{r}
we <- glm(fate ~ food, data = binomWE, family = "binomial")

summary(we)

anova(we, test = "Chisq")
```

# Growth trajectories of different treatment groups

## Manipulate the data
```{r}
# Internal process of the loop to convert days to dates
expt_day <- function(start.date, date){
  date <- mdy(date)
  start <- mdy(start.date)
  diff <- yday(date) - yday(start)

  return(diff)
}

# Code for the loop itself
loop <- function(start.date, date.list){
  days <- rep(NA, length(date.list))
  for (i in 1:(length(days))){
    days[i] <- expt_day(start.date = start.date, date = date.list[i])
  }
  return(days)
}
```

```{r}
data$day.hatch <- loop(start.date = "07/12/2020", date.list = data$date.hatch)
data$day.3rd <- loop(start.date = "07/12/2020", date.list = data$date.3rd)
data$day.4th <- loop(start.date = "07/12/2020", date.list = data$date.4th)
data$day.5th <- loop(start.date = "07/12/2020", date.list = data$date.5th)
data$day.w <- loop(start.date = "07/12/2020", date.list = data$date.w)

data$day.5th.2 <- loop(start.date = "07/12/2020", date.list = data$date.5th.2)
data$day.5th.3 <- loop(start.date = "07/12/2020", date.list = data$date.5th.3)
data$day.6th <- loop(start.date = "07/12/2020", date.list = data$date.6th)
data$day.6th.2 <- loop(start.date = "07/12/2020", date.list = data$date.6th.2)
data$day.6th.3 <- loop(start.date = "07/12/2020", date.list = data$date.6th.3)
data$day.7th <- loop(start.date = "07/12/2020", date.list = data$date.7th)
data$day.7th.2 <- loop(start.date = "07/12/2020", date.list = data$date.7th.2)
data$day.7th.3 <- loop(start.date = "07/12/2020", date.list = data$date.7th.3)
data$day.coc <- loop(start.date = "07/12/2020", date.list = data$date.coc)

data$tthatch <- 0
data$tt3rd <- data$day.3rd - data$day.hatch
data$tt4th <- data$day.4th - data$day.hatch
data$tt5th <- data$day.5th - data$day.hatch
data$ttw <- data$day.w - data$day.hatch

data$tt5th.2 <- data$day.5th.2 - data$day.hatch
data$tt5th.3 <- data$day.5th.3 - data$day.hatch
data$tt6th <- data$day.6th - data$day.hatch
data$tt6th.2 <- data$day.6th.2 - data$day.hatch
data$tt6th.3 <- data$day.6th.3 - data$day.hatch
data$tt7th <- data$day.7th - data$day.hatch
data$tt7th.2 <- data$day.7th.2 - data$day.hatch
data$tt7th.3 <- data$day.7th.3 - data$day.hatch
data$ttcoc <- data$day.coc - data$day.hatch

```

```{r}
data$ln.hatch <- log(data$mass.hatch)
data$ln.3rd <- log(data$mass.3rd)
data$ln.4th <- log(data$mass.4th)
data$ln.5th <- log(data$mass.5th)
data$ln.w <- log(data$mass.w)

data$ln.5th.2 <- log(data$mass.5th.2)
data$ln.5th.3 <- log(data$mass.5th.3)
data$ln.6th <- log(data$mass.6th)
data$ln.6th.2 <- log(data$mass.6th.2)
data$ln.6th.3 <- log(data$mass.6th.3)
data$ln.7th <- log(data$mass.7th)
data$ln.7th.2 <- log(data$mass.7th.2)
data$ln.7th.3 <- log(data$mass.7th.3)
data$ln.coc <- log(data$mass.remove)

```

```{r}
dataslim <- data[data$treatment == "NP NHS" | data$treatment == "NP HS", ] %>%
  select(ID, food, fate, treatment, tthatch, ln.hatch, tt3rd, ln.3rd, tt4th, ln.4th, tt5th, ln.5th, ttw, ln.w)

dataslim2 <- data[data$treatment == "P NHS" | data$treatment == "P HS", ] %>%
  select(ID, food, fate, treatment, tthatch, ln.hatch, tt3rd, ln.3rd, tt4th, ln.4th, tt5th, ln.5th, tt5th.2, 
         ln.5th.2, tt5th.3, ln.5th.3, tt6th, ln.6th, tt6th.2, ln.6th.2, tt6th.3, ln.6th.3, tt7th, ln.7th, 
         tt7th.2, ln.7th.2, tt7th.3, ln.7th.3, ttcoc, ln.coc )
```

## NP NHS vs NP HS

### More manupulation

```{r}
masslong <- dataslim %>%
  gather(instar, mass, ln.hatch, ln.3rd, ln.4th, ln.5th, ln.w)
masslong$instar <- gsub("ln.", "", masslong$instar)

agelong <- dataslim %>%
  gather(instar, age, tthatch, tt3rd, tt4th, tt5th, ttw)
agelong$instar <- gsub("tt", "", agelong$instar)

masslong <- masslong %>% 
  select(ID, treatment, food, instar, mass)
agelong <- agelong %>% 
  select(ID, treatment, food, instar, age)

datalong<-merge(masslong, agelong, by=c("ID", "treatment", "food", "instar"))

datalong$instar <- factor(datalong$instar, levels = c("hatch", "3rd", "4th", "5th", "w"))

datalong$newtreatment <- paste(datalong$treatment, datalong$food)
datalong$newtreatment <- factor(datalong$newtreatment, levels = c("NP NHS DC", "NP NHS TB", "NP HS DC", "NP HS TB"))
```

```{r}

#calculate means and variation

#constructing summarySE for age and mass to make mean plot

masssum <- summarySE(datalong, measurevar = "mass",
                  groupvars = c("newtreatment", "instar"),
                  na.rm=TRUE)
masssum

agesum <- summarySE(datalong, measurevar = "age",
                   groupvars = c("newtreatment", "instar"),
                   na.rm = TRUE)
agesum

#combine into one data frame

masssum$age <- agesum[, 4]
masssum$age.se <- agesum[, 6]
masssum$age.ci <- agesum[, 7]
```

### Plot the data (with CIs)

```{r}
meanplot <- ggplot(masssum, aes(x = age, y = mass, group = newtreatment, color = newtreatment, shape= newtreatment)) + 
  theme_bw() + 
  geom_point(size = 4) + 
  geom_line(aes(linetype = newtreatment), size = 1) + 
  geom_errorbar(aes(ymin = mass-(ci/2), ymax = mass + (ci/2)),
                width = .3, size = 1) + 
  geom_errorbarh(aes(xmin = age - (age.ci/2), xmax = age + (age.ci/2)),
                 height = .3, size = 1) + 
  scale_color_manual(values = c("#D81B60","#1E88E5", "#D81B60","#1E88E5" )) +
  scale_shape_manual(values = c(1, 1, 2, 2)) + 
  scale_linetype_manual(values = c("dashed", "dashed", "solid", "solid")) + 
  labs(x = "Total developmental age (days)", y = "Total mass gained (ln)")

ggsave("meanplot.png", plot = meanplot, width = 8, height = 5)

meanplot
```

## P NHS v P HS

### Data manipulation - log scale

```{r}
masslong2 <- dataslim2 %>%
  gather(instar, mass, ln.hatch, ln.3rd, ln.4th, ln.5th, ln.5th.2, ln.5th.3, ln.6th, ln.6th.2, ln.6th.3, ln.7th,
         ln.7th.2, ln.7th.3, ln.coc)
masslong2$instar <- gsub("ln.", "", masslong2$instar)

agelong2 <- dataslim2 %>%
  gather(instar, age, tthatch, tt3rd, tt4th, tt5th, tt5th.2, tt5th.3, tt6th, tt6th.2, tt6th.3, tt7th, tt7th.2,
         tt7th.3, ttcoc)
agelong2$instar <- gsub("tt", "", agelong2$instar)

masslong2 <- masslong2 %>% 
  select(ID, treatment, food, instar, mass)
agelong2 <- agelong2 %>% 
  select(ID, treatment, food, instar, age)

datalong2<-merge(masslong2, agelong2, by=c("ID", "treatment", "food", "instar"))

#datalong2$instar <- factor(datalong2$instar, levels = c("hatch", "3rd", "4th", "5th", "w"))

datalong2$newtreatment <- paste(datalong2$treatment, datalong2$food)
datalong2$newtreatment <- factor(datalong2$newtreatment, levels = c("P NHS DC", "P NHS TB", "P HS DC", "P HS TB"))
```

### Individual trajectories - log scale

```{r}
spaghettiplot <- ggplot(datalong2, aes(x = age, y = mass, group = ID, color = newtreatment, 
                                     shape= newtreatment)) + 
  theme_bw() + 
  geom_point() + 
  geom_line(aes(linetype = newtreatment), size = .8) + 
  #geom_errorbar(aes(ymin = mass-(ci/2), ymax = mass + (ci/2)),
               # width = .3, size = 1) + 
  #geom_errorbarh(aes(xmin = age - (age.ci/2), xmax = age + (age.ci/2)),
               #  height = .3, size = 1) + 
  scale_color_manual(values = c("#D81B60","#1E88E5", "#D81B60","#1E88E5" )) +
  scale_shape_manual(values = c(1, 1, 2, 2)) + 
  scale_linetype_manual(values = c("solid", "solid", "dashed", "dashed")) + 
  labs(x = "Total developmental age (days)", y = "Total mass gained (ln)")

spaghettiplot

ggsave("~/Desktop/lnspaghettiplot.png", plot = spaghettiplot, width = 8, height = 5)
```

### Data manipulation - normal scale

```{r}
dataslim3 <- data[data$treatment == "P NHS" | data$treatment == "P HS", ] %>%
  select(ID, food, fate, treatment, tthatch, mass.hatch, tt3rd, mass.3rd, tt4th, mass.4th, tt5th, mass.5th, tt5th.2, 
         mass.5th.2, tt5th.3, mass.5th.3, tt6th, mass.6th, tt6th.2, mass.6th.2, tt6th.3, mass.6th.3, tt7th, mass.7th, 
         tt7th.2, mass.7th.2, tt7th.3, mass.7th.3, ttcoc, mass.remove )

dataslim3$mass.coc <- dataslim3$mass.remove

masslong3 <- dataslim3 %>%
  gather(instar, mass, mass.hatch, mass.3rd, mass.4th, mass.5th, mass.5th.2, mass.5th.3, mass.6th, mass.6th.2,
         mass.6th.3, mass.7th, mass.7th.2, mass.7th.3, mass.coc)
masslong3$instar <- gsub("mass.", "", masslong3$instar)

agelong3 <- dataslim3 %>%
  gather(instar, age, tthatch, tt3rd, tt4th, tt5th, tt5th.2, tt5th.3, tt6th, tt6th.2, tt6th.3, tt7th, tt7th.2,
         tt7th.3, ttcoc)
agelong3$instar <- gsub("tt", "", agelong2$instar)

masslong3 <- masslong3 %>% 
  select(ID, treatment, food, instar, mass)
agelong3 <- agelong3 %>% 
  select(ID, treatment, food, instar, age)

datalong3<-merge(masslong3, agelong3, by=c("ID", "treatment", "food", "instar"))

#datalong2$instar <- factor(datalong2$instar, levels = c("hatch", "3rd", "4th", "5th", "w"))

datalong3$newtreatment <- paste(datalong3$treatment, datalong3$food)
datalong3$newtreatment <- factor(datalong3$newtreatment, levels = c("P NHS DC", "P NHS TB", "P HS DC", "P HS TB"))
```

### Graphing - normal scale

```{r}
spaghettiplot2 <- ggplot(datalong3, aes(x = age, y = mass, group = ID, color = newtreatment, 
                                     shape= newtreatment)) + 
  theme_bw() + 
  geom_point() + 
  ylim(0, 20000) + 
  geom_line(aes(linetype = newtreatment), size = .8) + 
  #geom_errorbar(aes(ymin = mass-(ci/2), ymax = mass + (ci/2)),
               # width = .3, size = 1) + 
  #geom_errorbarh(aes(xmin = age - (age.ci/2), xmax = age + (age.ci/2)),
               #  height = .3, size = 1) + 
  scale_color_manual(values = c("#D81B60","#1E88E5", "#D81B60","#1E88E5" )) +
  scale_shape_manual(values = c(1, 1, 2, 2)) + 
  scale_linetype_manual(values = c("solid", "solid", "dashed", "dashed")) + 
  labs(x = "Total developmental age (days)", y = "Total mass gained (mg)")

spaghettiplot2

ggsave("~/Desktop/spaghettiplot.png", plot = spaghettiplot2, width = 8, height = 5)
```

# Wasp emergence

We need to make sure the numbers of wasps emerging and eclosing from TB PNHS and DC PNHS caterpillars is the same with this data.

## Manipulate the data

```{r}
datawasp <- data %>%
  select(ID, food, treatment, date.coc, mass.remove, wet, num.w.coc, num.w.larvae, num.w.eclos)

datawasp <- datawasp[complete.cases(datawasp$num.w.coc), ]
```

```{r}
datawasp$num.w.tot <- datawasp$num.w.coc + datawasp$num.w.larvae
datawasp$prop.eclos <- datawasp$num.w.eclos / datawasp$num.w.tot
```

## Start graphing and running stats

```{r}
ggplot(data = datawasp, aes(x = treatment, y = num.w.tot)) + 
  theme_bw() + 
  geom_boxplot()
```

```{r}
ggplot(data = datawasp, aes(x = food, y = num.w.tot)) + 
  theme_bw() + 
  geom_boxplot() + 
  labs(x = "Food Type", y = "Total number of wasps emerged")
```

```{r}
waspmod1 <- lm(num.w.tot ~ food, data = datawasp[datawasp$treatment == "P NHS", ])

anova(waspmod1)
```

```{r}
ggplot(data = datawasp[datawasp$treatment == "P NHS", ], aes(x = food, y = prop.eclos)) + 
  theme_bw() + 
  geom_boxplot() + 
  labs(x = "Food Type", y = "Proportion of emerged wasps successfully eclosed")
```

```{r}
waspmod2 <- glm(prop.eclos ~ food, family = "binomial", data = datawasp[datawasp$treatment == "P NHS", ])

summary(waspmod2)
```

```{r}
ggplot(data = datawasp, aes(x = mass.remove, y = num.w.tot, color = food)) + 
  theme_bw() + 
  geom_point()
```

# Redone development time plot - means and SDs

```{r}
data$day.max <- loop(start.date = "07/12/2020", date.list = data$date.max)
data$ttmax <- data$day.max - data$day.hatch
```

```{r}
datalong4 <- data %>%
  select(ID, food, treatment, mass.max, ttmax)

meantime <- summarySE(datalong4, measurevar = "ttmax", groupvars = c("food", "treatment"), na.rm = TRUE)

meanmass <- summarySE(datalong4, measurevar = "mass.max", groupvars = c("food", "treatment"), na.rm = TRUE)

datameans <- meantime %>%
  select(food, treatment, ttmax, sd)

datameans$mass.max <- meanmass$mass.max
datameans$sdmass <- meanmass$sd
```

```{r}
meansbytrt <- ggplot(aes(x = ttmax, y = mass.max, color = food, shape = treatment), data = datalong4) +
  theme_bw() + 
  xlim(15, 45) + 
  geom_point(alpha = 0.4) + 
  scale_shape_manual(values = c(19, 1, 17, 2)) + 
  labs(x = "Development time at maximum mass (days)", y = "Maximum mass achieved (g)", 
       shape = "Treatment", color = "Hostplant Consumed") + 
  geom_point(data = datameans, aes(x = ttmax, y = mass.max, color = food, shape = treatment), size = 4) + 
  geom_errorbar(data = datameans, aes(ymin = mass.max - sdmass, ymax = mass.max + sdmass), size = .5) + 
  geom_errorbarh(data = datameans, aes(xmin = ttmax - sd, xmax = ttmax + sd), size = .5) + 
  guides(size = FALSE)

meansbytrt

ggsave("~/Desktop/meansbytrt.png", plot = meansbytrt, width = 8, height = 5)
```

## Add points onto developmental outcomes graph

```{r}
datalong5 <- data %>%
  select(ID, food, fate, mass.max, ttmax, area)

meantime2 <- summarySE(datalong5, measurevar = "ttmax", groupvars = c("food", "fate"), na.rm = TRUE)

meanmass2 <- summarySE(datalong5, measurevar = "mass.max", groupvars = c("food", "fate"), na.rm = TRUE)
```

### Group them into one dataframe

```{r}
datameans2 <- meantime2 %>%
  select(food, fate, ttmax, se)

datameans2$mass.max <- meanmass2$mass.max
datameans2$semass <- meanmass2$se

```

## Graphs with means only

### Maximum mass gained and age at that point