### 2023-01-03 CodeForSubmission.Rmd
Created complete script for manuscript submission

---
title: "Code for Submission"
output: html_notebook
---

Authors: Anna Parker and Joel Kingsolver

This code contains all analyses and figures for the manuscript. It is organized in order of appearance in the manuscript, with the necessary data manipulations happening between steps as required.

Table of Contents: 
0. Load in data and packages
1. Initial data manipulation
2. Figures, tables, and statistics (in order of appearance)
  a. Figure 1
  b. Figure 2 and associated stats
  c. Figure 3 and Table 1 and associated stats
  d. Figure 4, Figure 5, Table 2, Supplemental Table 2, and associated stats

## 0. Load in data and packages

```{r}
library(tidyverse)
library(ggplot2)
library(Rmisc)
library(lubridate)
library(lmerTest)
library(nnet)
library(jmv)
```

```{r}
data <- read.csv("~/Desktop/MsCcHSFeedJan1MOD.csv")
scan <- read.csv("~/Desktop/MsCcHSFeedLeafScanFinal.csv", header = T)
datadead <- read.csv("~/Desktop/MsCcHSFeedJan1MOD.csv")
```

## 1. Initial data manipulation

```{r}
data <- data[data$fate != "dead" & data$fate != "quink" & data$fate != "wander*" & data$fate != "?", ]
data <- data[complete.cases(data[, 1]), ]
data$fate <- factor(data$fate, levels = c("wander", "em", "wowe", "wowe mongo"))

```

```{r}
datadead <- datadead[datadead$fate != "quink" & datadead$fate != "wander*" & datadead$fate != "?", ]
datadead <- datadead[complete.cases(datadead[, 1]), ]
datadead$treatment <- factor(datadead$treatment, levels = c("NP NHS", "NP HS", "P NHS", "P HS"))
```

## 2. Figures and statistics

### a. Figure 1

```{r}
data$treatment <- factor(data$treatment, levels = c("NP NHS", "NP HS", "P NHS", "P HS"))

dataviolin <- data
dataviolin$fate <- gsub(" mongo", "", dataviolin$fate)
dataviolin$fate <- factor(dataviolin$fate, levels = c("wander", "em", "wowe"))
```

#### Graph

```{r}
violins <- ggplot(data = dataviolin, aes(x = treatment, y = mass.max)) + 
  geom_violin(fill = "light grey", color = "light grey") + 
  geom_jitter(aes(color = fate, shape = food), width = 0.1, height = 0) + 
  theme_bw() + 
  labs(x = "Treatment Group", y  = "Maximum mass achieved (mg)",  color = "Developmental Outcome", shape = "Hostplant") + 
  scale_color_manual(values = c("#1E88E5", "lime green", "#D81B60"), 
                     labels = c("Wander", "WE", "WOWE")) + 
  scale_shape_manual(values = c(1, 19))

violins
```

*Brackets and text annotated in Photoshop*

### b. Figure 2 and stats

```{r}
datadead$treatment <- factor(datadead$treatment, levels = c("NP NHS", "NP HS", "P NHS", "P HS"))
datadead$fate <- factor(datadead$fate, levels = c("dead", "wander", "em", "wowe", "wowe mongo"))
```

**Make dataframe for annotating in the sample sizes**

```{r}
datadeadsum <- summarySE(datadead, 
                         measurevar = "mass.max", #doesn't matter, just need the Ns, will be NAs anyways
                         groupvars = c("treatment", "food"))

samplesize <- data.frame(N = datadeadsum$N, 
                         treatment = datadeadsum$treatment,
                         food = datadeadsum$food, 
                         x = c(1, 1, 2, 2, 3, 3, 4, 4), 
                         y = 1.05,
                         fate = NA)
```

#### Graph

```{r}
barsdead <- ggplot(data = datadead, aes(x = treatment, fill = fate)) + 
  theme_bw() + 
  geom_bar(position = "fill") + 
  facet_wrap(~food) + 
  scale_y_continuous(breaks = c(0, .25, .5, .75, 1), limits = c(0, 1.07)) + 
  labs(x = "Treatment group", y = "Proportion of indiviudals", fill = "Developmental Outcome") + 
  scale_fill_manual(values = c("dead" = "purple", "wander" = "#1E88E5" , "em" = "lime green", 
                                       "wowe" = "orange", "wowe mongo" =  "#D81B60"),
                    labels = c("Died", "Wander", "WE", "Small WOWE", "Large WOWE")) + 
  geom_text(data = samplesize, aes(x = x, y = y, label = N))

barsdead
```

*Note: Could not get the legend to stop labelling in alphabetical order, so I will switch colors in Photoshop :( *

#### Stats

Multinomial regression to report proportions of outcomes in PHS treatment group

```{r}
datamulti <- data[data$para == "yes" & data$HS == "yes", ]
datamulti$fate <- factor(datamulti$fate, levels = c("em", "wander", "wowe", "wowe mongo"), labels = c("WE", "Wander", "Small WOWE", "Large WOWE"))
datamulti$food <- factor(datamulti$food, levels = c("TB", "DC"))
```

Use the multinom command to run a model

```{r}
fatemodel <- multinom(fate ~ food, data = datamulti)

summary(fatemodel)
```

Get p-values:

```{r}
fatez <- summary(fatemodel)$coefficients/summary(fatemodel)$standard.errors
#2-tailed z test
pfate <- (1-pnorm(abs(fatez), 0, 1))*2
pfate
```

### c. Figure 3, Table 1, Stats

```{r}
datawasp <- data[data$para == "yes", ] %>%
  select(ID, food, treatment, num.w.coc, num.w.larvae, num.w.eclos)

datawasp <- datawasp[complete.cases(datawasp), ]

datawasp$num.w.tot <- datawasp$num.w.coc + datawasp$num.w.larvae
datawasp$percent <- (datawasp$num.w.eclos) / (datawasp$num.w.tot)
datawasp$ln.w.tot <- log(datawasp$num.w.tot + 1)
```

#### Table

```{r}
eclosmeans <- summarySE(datawasp, measurevar = "percent", groupvars = c("food", "treatment"))

eclosmeans
```

More manipulation to graph...

```{r}
datawasp2 <- datawasp %>%
  gather(stage, number, num.w.tot, num.w.eclos)

datawasp2$stage <- gsub("num.w.", "", datawasp2$stage)
datawasp2$stage <- gsub("tot", "Emerged", datawasp2$stage)
datawasp2$stage <- gsub("eclos", "Eclosed", datawasp2$stage)

datawasp2 <- datawasp2 %>%
  select(ID, food, treatment, stage, number)

waspmeans <- summarySE(datawasp2, measurevar = "number", groupvars = c("food", "treatment", "stage"))
waspmeans$stage <- factor(waspmeans$stage, levels = c("Emerged", "Eclosed"))
waspmeans$comb <- paste(waspmeans$food, waspmeans$treatment, sep = " ")
```

#### Graph

```{r}
newwaspplot <- ggplot(data = waspmeans, aes(x = stage, y = number, group = comb, color = treatment)) + 
  theme_bw() + 
  geom_line(aes(linetype = food)) + 
  geom_point(aes(shape = food), size = 3) + 
  geom_errorbar(aes(ymin = number - se, ymax = number + se), size = 0.25, width = 0.2) + 
  scale_linetype_manual(values = c("dashed", "solid")) + 
  scale_shape_manual(values = c(1, 19)) + 
  labs(x = "Wasp Developmental Stage", y = "Number of wasps per caterpillar host", shape = "Hostplant", 
       linetype = "Hostplant", color = "Treatment")

newwaspplot
```

#### Stats

Eclosion success:
```{r}
waspbinom <- glm(percent ~ food*treatment, family = "binomial", weights = num.w.tot, data = datawasp)

#summary(waspbinom)

anova(waspbinom, test = "Chisq")

```

Emergence totals:
```{r}
waspemerge <- lm(ln.w.tot ~ food*treatment, data = datawasp)

#summary(waspemerge)

anova(waspemerge)
```

### d. Figure 4 + 5, Table 2 + Supp Table 2, Stats

#### Data manipulation
```{r}
# Internal process of the loop to convert days to dates
expt_day <- function(start.date, date){
  date <- mdy(date)
  start <- mdy(start.date)
  diff <- yday(date) - yday(start)

  return(diff)
}

# Code for the loop itself
loop <- function(start.date, date.list){
  days <- rep(NA, length(date.list))
  for (i in 1:(length(days))){
    days[i] <- expt_day(start.date = start.date, date = date.list[i])
  }
  return(days)
}
```

```{r}
data$day.hatch <- loop(start.date = "07/12/2020", date.list = data$date.hatch)
data$day.max <- loop(start.date = "07/12/2020", date.list = data$date.max)
data$ttmax <- data$day.max - data$day.hatch

#Convert from mg to g
data$mass.max <- data$mass.max / 1000
```

Code to get total leaf area consumed:

```{r}
#0 - make unique ID for each caterpillar and leaf eaten
scan$comb <- paste(scan$Caterpillar.ID, "_", scan$Leaf.ID)

#1 - make an output file
unique <- unique(scan$comb)
consumed <- as.data.frame(unique)
consumed$area <- 0

#2-4 - loop through and subset by each unique ID (before and after), then subtract to get area
for (i in 1:length(consumed$unique)){
    subset <- filter(scan, scan$comb == consumed[i, 1])
    value <- max(subset$Leaf.Area) - min(subset$Leaf.Area)
    consumed[i, 2] <- value
}

#5 - reseparate cat and leaf ID
consumed <- consumed %>%
  separate(unique, c("catID", "leafID"), "_")

#6 - make a new output file
ID <- unique(consumed$catID)
consumedsum <- as.data.frame(ID)
consumedsum$totalarea <- 0

#7 - loop through and subset by cat ID and add all areas together
for (i in 1:length(consumedsum$ID)){
  subset2 <- filter(consumed, consumed$catID == consumedsum[i, 1])
  value2 <- sum(subset2$area)
  consumedsum[i, 2] <- value2
}

#8 - final data wrangling 
consumedsum$ID <- as.numeric(consumedsum$ID)

consumedsum <- consumedsum[consumedsum$ID %in% data$ID, ]

data <- data[data$ID %in% consumedsum$ID, ]

consumedsum <- consumedsum[order(consumedsum$ID), ]

#9 - add to dataframe
data$area <- consumedsum$totalarea

#10 - change to m^2 fromm cm^2
data$area <- data$area / 10000

```

Make new dataframes
```{r}
datalong5 <- data %>%
  select(ID, food, fate, mass.max, ttmax, area)

meantime2 <- summarySE(datalong5, measurevar = "ttmax", groupvars = c("food", "fate"), na.rm = TRUE)

meanmass2 <- summarySE(datalong5, measurevar = "mass.max", groupvars = c("food", "fate"), na.rm = TRUE)

meanarea2 <- summarySE(datalong5, measurevar = "area", groupvars = c("food", "fate"), na.rm = TRUE)
```

Group means together

```{r}
datameans2 <- meantime2 %>%
  select(food, fate, ttmax, se)

datameans2$mass.max <- meanmass2$mass.max
datameans2$semass <- meanmass2$se
datameans2$area <- meanarea2$area
datameans2$arease <- meanarea2$se

```

#### Graph Fig 4

```{r}
newmeans <- ggplot(data = datalong5, aes(x = ttmax, y = mass.max, color = fate, shape = food)) + 
  theme_bw() + 
  xlim(15, 45) +
  geom_point(alpha = 0.4) + 
  scale_shape_manual(values = c(1, 19)) + 
  scale_color_manual(values = c("#1E88E5", "lime green", "orange", "#D81B60"), 
                     labels = c("Wander", "WE", "Small WOWE", "Large WOWE")) + 
  geom_point(aes(x = ttmax, y = mass.max, color = fate, shape = food), data = datameans2, size = 4) + 
  geom_errorbar(data = datameans2, aes(ymin = mass.max - semass, ymax = mass.max + semass), size = .5) + 
  geom_errorbarh(data = datameans2, aes(xmin = ttmax - se, xmax = ttmax + se), size = .5) + 
  guides(size = FALSE) + 
  labs(x = "Age at maximum mass (days)", y = "Maximum mass achieved (g)", 
       shape = "Hostplant Consumed", color = "Developmental Outcome")

newmeans
```

#### Graph Fig 5

```{r}
newarea <- ggplot(data = datalong5, aes(x = area, y = mass.max, color = fate, shape = food)) + 
  theme_bw() + 
  geom_point(alpha = 0.4) + 
  scale_shape_manual(values = c(1, 19)) + 
  scale_color_manual(values = c("#1E88E5", "lime green", "orange", "#D81B60"), 
                     labels = c("Wander", "WE", "Small WOWE", "Large WOWE")) + 
  ylim(0, 20) + 
  xlim(0, .6) +
  geom_point(aes(x = area, y = mass.max, color = fate, shape = food), data = datameans2, size = 4) + 
  geom_errorbar(data = datameans2, aes(ymin = mass.max - semass, ymax = mass.max + semass), size = .5, width = 0) + 
  geom_errorbarh(data = datameans2, aes(xmin = area - arease, xmax = area + arease), size = .5) + 
  guides(size = FALSE) + 
  labs(x = expression(paste("Hostplant leaf area consumed ", (m^{2}))), y = "Maximum mass achieved (g)", 
       shape = "Hostplant Consumed", color = "Developmental Outcome")

newarea
```

#### Stats

*Each tests developmental outcome vs. treatment, then compares them via AIC*

**Maximum age differences across treatment and outcome**

```{r}
maxage <- lmer(ttmax ~ fate*food + (1|bin), data = data)

anova(maxage)
```

```{r}
maxage2 <- lmer(ttmax ~ para*HS*food + (1|bin), data = data)

anova(maxage2)
```

```{r}
AIC(maxage, maxage2)
```

**Maximum mass**

```{r}
maxmass <- lmer(mass.max ~ fate*food + (1|bin), data = data)

anova(maxmass)
```

```{r}
maxmass2 <- lmer(mass.max ~ para*HS*food + (1|bin), data = data)

anova(maxmass2)
```

```{r}
AIC(maxmass, maxmass2)
```

**Leaf area consumed**

```{r}
areatrt <- lmer(area ~ fate*food + (1|bin), data = data)

anova(areatrt)
```

```{r}
areatrt2 <- lmer(area ~ para*HS*food + (1|bin), data = data)

anova(areatrt2)
```

```{r}
AIC(areatrt, areatrt2)
```